---
title: "PBMC_Multimodal_dataset_Ng_Turiello"
author: "Dillon Corvino"
date: "30/03/2021"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    highlight: tango
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Built with R version `r getRversion()`

## Setup {.tabset}

### Dataset information
```{r Dataset_Info}

# CITEseq reference of 162,000 PBMCs measured with 228 antibodies 

# PBMCS
# 8 donors
# 3 timepoints - Day 0, 3, 7 
# Treatment = vaccination with HIV vaccine

```

### Environment
```{r setup}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,  
  eval = TRUE, 
  tidy = TRUE
)

knitr::opts_knit$set(
  root.dir = "../"
)


# Environment Set up
rm(list = ls()) #Clean workspace
cat("\014")     #Clean Console
gc() # Free memory

# pipeline variables
long.compute <- FALSE

# Establish working directory
library("rstudioapi")

# Set working directory to source file location
setwd(dirname(getActiveDocumentContext()$path))
setwd("..")

# Create output directories & load custom functions & colour scheme
source("scripts/Setup.R", local = knitr::knit_global())

```

### Load dataset & filter dataset
```{r load_dataset}

if(long.compute){
  
  pbmc.seurat <- LoadH5Seurat("data/pbmc_multimodal.h5seurat")
  
  # 8 donors across three timepoints of pre, and 3 or 7 days post HIV vaccination
  table(pbmc.seurat@meta.data$time) # approx 50K cells per timepoint 
  table(pbmc.seurat@meta.data$time, pbmc.seurat@meta.data$donor) # 8 donors 
  
  # remove doublets 
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$celltype.l2
  
  pbmc.seurat <- subset(pbmc.seurat, idents = "Doublet", invert = TRUE)
  
  
  # set defaults
  DefaultAssay(pbmc.seurat) <- "SCT"
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$celltype.l1
  
  
  # create custom annotation of cell types 
  Bcells <- c("B intermediate lambda", "B naive kappa", "B intermediate kappa", "B memory kappa", "B naive lambda", "B memory lambda")
  plasma <- c("Plasma", "Plasmablast")
  nk.dim <- c(paste0("NK_", 1:4))
  treg <- c("Treg Naive", "Treg Memory")
  dnT <- c("dnT_1", "dnT_2")
  cDC2 <- c("cDC2_1", "cDC2_2", "ASDC_mDC")
  pDC <- c("ASDC_pDC", "pDC")
  gdT <- c("gdT_2", "gdT_3", "gdT_4") 
  cd4.cm <- c(paste0("CD4 TCM_", 1:3))
  cd4.em <- c(paste0("CD4 TEM_", 1:4))
  cd8.naive <- c("CD8 Naive", "CD8 Naive_2")
  cd8.em <- c(paste0("CD8 TEM_", 1:6))
  cd8.cm <- c(paste0("CD8 TCM_", 1:3))
  
  pbmc.seurat@meta.data <- pbmc.seurat@meta.data %>%
    dplyr::mutate(custom_anno = case_when(celltype.l3 %in% Bcells ~ "B_cells", 
                                          celltype.l3 %in% plasma ~ "Plasmablasts", 
                                          celltype.l3 %in% nk.dim ~ "NK_Dim", 
                                          celltype.l3 %in% treg ~ "Treg", 
                                          celltype.l3 %in% dnT ~ "dnT", 
                                          celltype.l3 %in% cDC2 ~ "cDC2", 
                                          celltype.l3 %in% pDC ~ "pDC", 
                                          celltype.l3 %in% gdT ~ "gdT", 
                                          celltype.l3 %in% cd4.cm ~ "CD4_CM", 
                                          celltype.l3 %in% cd4.em ~ "CD4_EM", 
                                          celltype.l3 %in% cd8.naive ~ "CD8_Naive", 
                                          celltype.l3 %in% cd8.em ~ "CD8_EM", 
                                          celltype.l3 %in% cd8.cm ~ "CD8_CM", 
                                          celltype.l3 == "gdT_1" ~ "gdT_V9D2",
                                          TRUE ~ celltype.l3))
  
  # additional formatting of custom_anno
  pbmc.seurat@meta.data$custom_anno <- gsub(" ", "_", pbmc.seurat@meta.data$custom_anno)
  pbmc.seurat@meta.data$custom_anno <- gsub("^Platelet$", "Platelets", pbmc.seurat@meta.data$custom_anno)
  pbmc.seurat@meta.data$custom_anno <- gsub("Proliferating", "Prolif", pbmc.seurat@meta.data$custom_anno)
  pbmc.seurat@meta.data$custom_anno <- gsub("^NK_CD56bright$", "NK_Bright", pbmc.seurat@meta.data$custom_anno)
  
  
  # set levels for factor 
  cells.group.1 <- c("cDC1", "cDC2", "pDC", "CD14_Mono", "CD16_Mono", "Platelets", "Eryth", "HSPC", "B_cells", "Plasmablasts")
  cells.group.2 <- c("CD4_Naive", "CD4_CM", "CD4_EM", "CD4_CTL", "Treg")
  cells.group.3 <- c("dnT",  "CD8_Naive", "CD8_CM", "CD8_EM")
  cells.group.4 <- c("gdT", "gdT_V9D2", "MAIT")
  cells.group.5 <- c("NK_Prolif", "CD8_Prolif", "CD4_Prolif")
  cells.group.6 <- c("ILC", "NK_Bright", "NK_Dim")
  
  cell.groups <- list("Myeloid" = cells.group.1, 
                      "CD4" = cells.group.2, 
                      "CD8" = cells.group.3,
                      "Innate_like" = cells.group.4, 
                      "Prolif" = cells.group.5, 
                      "Innate" = cells.group.6)
  
  pbmc.seurat@meta.data$custom_anno <- factor(pbmc.seurat@meta.data$custom_anno, levels = c(cell.groups$Myeloid, 
                                                                                            cell.groups$CD4, 
                                                                                            cell.groups$CD8, 
                                                                                            cell.groups$Innate_like, 
                                                                                            cell.groups$Prolif, 
                                                                                            cell.groups$Innate))
  
  # save rds object cell.groups 
  saveRDS(cell.groups, "saves/cell_groups.rds")
  
  
  # amend "celltype.l1" annotation formatting
    pbmc.seurat@meta.data$celltype.l1 <- gsub(" ", "_", pbmc.seurat@meta.data$celltype.l1)

 
  # save seurat 
    #SaveH5Seurat(pbmc.seurat, "saves/pbmc_seurat_imputed.h5Seurat", overwrite = T)
}else{
  
  pbmc.seurat <- LoadH5Seurat("saves/pbmc_seurat_imputed.h5Seurat")
  
  # amend "celltype.l1" annotation formatting
  pbmc.seurat@meta.data$celltype.l1 <- gsub(" ", "_", pbmc.seurat@meta.data$celltype.l1)
  
  #set defaults
  DefaultAssay(pbmc.seurat) <- "SCT"
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  
  cell.groups <- readRDS("saves/cell_groups.rds")
  
  
}

```

## Dataset overview {.tabset}

### Broad visualisation of dataset
```{r Visualisation_of_PBMC_reference}

output.dir <- "results/figures/WNN_visualisations/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


# A few levels of data annotation exist 
vars.plot <- c("celltype.l1", "celltype.l2", "celltype.l3", "custom_anno", "donor", "Phase", "time")

for(i in seq_along(vars.plot)){
  
  print(
    scCustomize::DimPlot_scCustom(pbmc.seurat, 
                                  reduction = "wnn.umap",
                                  group.by = vars.plot[i],
                                  label = TRUE, 
                                  label.size = 3, 
                                  pt.size = 3,
                                  raster = T,
                                  raster.dpi = c(2400, 2400),
                                  repel = TRUE) + 
      NoLegend()
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "wnnUMAP_", name.var, ".pdf"))
  dev.off()
  
}


```

### Visualise Cytotoxic molecule expression
```{r vis_cytotoxic_across_whole_dataset}

DefaultAssay(pbmc.seurat) <- "SCT"

cts.genes <- grep("CTS.*", rownames(pbmc.seurat), value = T)
goi <- c(cytotoxicity.markers, cts.genes)



######################
# plot featureplots
######################

output.dir <- "results/figures/FeaturePlots/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


for(i in seq_along(goi)){
  print(
    FeaturePlot(pbmc.seurat,
                features = goi[i],
                reduction = "wnn.umap",
                order = TRUE, 
                pt.size = 1, 
                raster = TRUE) + 
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "wnnUMAP_FeaturePlot_", goi[i], ".pdf"))
  dev.off()
  
}


######################
# plot density plots
######################

output.dir <- "results/figures/DensityPlots/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

for(i in seq_along(goi)){
  print(
    scCustomize::Plot_Density_Custom(pbmc.seurat,
                                     features = goi[i],
                                     custom_palette = batlow.pal,
                                     reduction = "wnn.umap", 
                                     pt.size = 1) + 
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "wnnUMAP_DensityPlot_", goi[i], ".pdf"))
  dev.off()
  
}


######################
# Violin Plots
######################
output.dir <- "results/figures/Stacked_VlnPlots/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

vars.plot <- c("celltype.l1", "celltype.l2", "celltype.l3", "custom_anno")


for(i in seq_along(vars.plot)){
  print(
    scCustomize::Stacked_VlnPlot(pbmc.seurat, 
                                 features = cytotoxicity.markers,
                                 pt.size = 0, 
                                 group.by = vars.plot[i],
                                 x_lab_rotate = TRUE)
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "Stacked_vlnplots_cytotoxicity_markers_", name.var, ".pdf"))
  dev.off()
}


scCustomize::Stacked_VlnPlot(pbmc.seurat, 
                             features = cts.genes,
                             pt.size = 0, 
                             group.by = "custom_anno",
                             x_lab_rotate = TRUE)

dev.copy(pdf, paste0(output.dir, "Stacked_vlnplots_CTS_genes.pdf"))
dev.off()






######################
# Heatmaps
######################

output.dir <- "results/figures/Heatmaps/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


vars.plot <- c("celltype.l1", "celltype.l2", "celltype.l3", "custom_anno")

for(i in seq_along(vars.plot)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data[[vars.plot[i]]]
  
  average.seurat <- AverageExpression(pbmc.seurat, return.seurat = TRUE)
  
  print(
    DoHeatmap(average.seurat, 
              features = cytotoxicity.markers,
              draw.lines = FALSE, 
              raster = FALSE) + 
      scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "Heatmap_cytotoxicity_markers_", name.var, ".pdf"))
  dev.off()
  
  
  print(
    DoHeatmap(average.seurat, 
              features = cts.genes,
              draw.lines = FALSE, 
              raster = FALSE) + 
      scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "Heatmap_CTS_genes_", name.var, ".pdf"))
  dev.off()
  
}

```

## Imputation {.tabset}

### Imputate gene expression
```{r imputation}

if(long.compute){
  
  output.dir <- "saves/"
  if(!dir.exists(output.dir)){
    dir.create(output.dir, 
               recursive = T)}
  
  
  # check that genes of interest are in variable genes list 
  pbmc.seurat # 20,957 genes across 161,159 cells
  
  DefaultAssay(pbmc.seurat) <- "SCT"
  
  var.features <- VariableFeatures(pbmc.seurat)
  
  # are genes of interest in variable genes list 
  x <- grep(paste0(all.markers, collapse = "|"), var.features, value = T)
  print(x)
  length(x)
  length(all.markers)
  
  # yes all genes of interst are in variable features
  
  pbmc.seurat <- SeuratWrappers::RunALRA(pbmc.seurat, 
                                         genes.use = var.features)
  
  # Export dataset for easy import
  SeuratDisk::SaveH5Seurat(pbmc.seurat, "saves/pbmc_seurat_imputed.h5Seurat", overwrite = TRUE)
  
  DefaultAssay(pbmc.seurat) <- "SCT"
  
}


```

## Visualise {.tabset}

### Visualise Cytotoxic molecule expression -- imputed
```{r vis_cytotoxic_across_whole_dataset_imputed}

# set assay 
DefaultAssay(pbmc.seurat) <- "alra"

cts.genes <- grep("CTS.*", rownames(pbmc.seurat), value = T)
goi <- c(cytotoxicity.markers, cts.genes)



######################
# plot featureplots
######################

output.dir <- "results/figures/FeaturePlots_imputed/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


for(i in seq_along(goi)){
  print(
    FeaturePlot(pbmc.seurat,
                features = goi[i],
                reduction = "wnn.umap",
                order = TRUE, 
                pt.size = 1, 
                raster = TRUE) + 
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "wnnUMAP_FeaturePlot_", goi[i], ".pdf"))
  dev.off()
  
}


######################
# plot density plots
######################

output.dir <- "results/figures/DensityPlots_imputed/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

for(i in seq_along(goi)){
  print(
    scCustomize::Plot_Density_Custom(pbmc.seurat,
                                     features = goi[i],
                                     custom_palette = batlow.pal,
                                     reduction = "wnn.umap", 
                                     pt.size = 1) + 
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "wnnUMAP_DensityPlot_", goi[i], ".pdf"))
  dev.off()
  
}


######################
# Violin Plots
######################
output.dir <- "results/figures/Stacked_VlnPlots_imputed/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

vars.plot <- c("celltype.l1", "celltype.l2", "celltype.l3", "custom_anno")


for(i in seq_along(vars.plot)){
  print(
    scCustomize::Stacked_VlnPlot(pbmc.seurat, 
                                 features = cytotoxicity.markers,
                                 pt.size = 0, 
                                 group.by = vars.plot[i],
                                 x_lab_rotate = TRUE)
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "Stacked_vlnplots_cytotoxicity_markers_", name.var, ".pdf"))
  dev.off()
}


scCustomize::Stacked_VlnPlot(pbmc.seurat, 
                             features = cts.genes,
                             pt.size = 0, 
                             group.by = "custom_anno",
                             x_lab_rotate = TRUE)

dev.copy(pdf, paste0(output.dir, "Stacked_vlnplots_CTS_genes.pdf"))
dev.off()






######################
# Heatmaps
######################

output.dir <- "results/figures/Heatmaps_imputed/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


vars.plot <- c("celltype.l1", "celltype.l2", "celltype.l3", "custom_anno")

for(i in seq_along(vars.plot)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data[[vars.plot[i]]]
  
  average.seurat <- AverageExpression(pbmc.seurat, return.seurat = TRUE)
  
  print(
    DoHeatmap(average.seurat, 
              features = cytotoxicity.markers,
              draw.lines = FALSE, 
              raster = FALSE) + 
      scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "Heatmap_cytotoxicity_markers_", name.var, ".pdf"))
  dev.off()
  
  
  print(
    DoHeatmap(average.seurat, 
              features = cts.genes,
              draw.lines = FALSE, 
              raster = FALSE) + 
      scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")
  )
  
  name.var <- gsub("\\.", "_", vars.plot[i])
  dev.copy(pdf, paste0(output.dir, "Heatmap_CTS_genes_", name.var, ".pdf"))
  dev.off()
  
}

```

### Subsetted vlnplots granzymes -- imputed
```{r vlnplots_granzymes_imputed}

# set assay 
DefaultAssay(pbmc.seurat) <- "alra"

goi <- c(cytotoxicity.markers)


######################
# Violin Plots
######################
output.dir <- "results/figures/Stacked_VlnPlots_imputed_fig_1/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


for(i in 1:length(cell.groups)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  temp.seurat <- subset(pbmc.seurat, idents = cell.groups[[i]])
  
  p1 <- scCustomize::Stacked_VlnPlot(temp.seurat, 
                                     features = goi,
                                     pt.size = 0, 
                                     group.by = "custom_anno",
                                     x_lab_rotate = 90) 
  
  for(j in 1:length(p1)){
    p1[[j]] <- p1[[j]] + theme(
      plot.margin = margin(t = 0,
                           r = 10,
                           b = 0,
                           l = 0, unit = "cm")
    )
  }
  
  print(p1)
  
  dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_",  names(cell.groups)[i], "group_cytotoxicity_imputed_.pdf"))
  dev.off()
}


```



### Subsetted vlnplots NKG7 vs TIA1 -- Raw values -- TIA1 not found in imputed
```{r vlnplots_NKG7_vs_TIA1_imputed}

# set assay 
DefaultAssay(pbmc.seurat) <- "SCT"

goi <- c("NKG7", "TIA1")

######################
# Violin Plots
######################
output.dir <- "results/figures/PBMC_Stacked_VlnPlots_SCT_TIA1_and_NKG7/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


for(i in 1:length(cell.groups)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  temp.seurat <- subset(pbmc.seurat, idents = cell.groups[[i]])
  
  p1 <- scCustomize::Stacked_VlnPlot(temp.seurat, 
                                     features = goi,
                                     pt.size = 0, 
                                     group.by = "custom_anno",
                                     x_lab_rotate = 90) 
  
  for(j in 1:length(p1)){
    p1[[j]] <- p1[[j]] + theme(
      plot.margin = margin(t = 0,
                           r = 10,
                           b = 0,
                           l = 0, unit = "cm")
    )
  }
  
  print(p1)
  
  dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_",  names(cell.groups)[i], "group_NKG7_vs_TIA1_SCT.pdf"))
  dev.off()
}





# across all annotation
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno


scCustomize::Stacked_VlnPlot(pbmc.seurat, 
                             features = goi,
                             pt.size = 0, 
                             group.by = "custom_anno",
                             x_lab_rotate = 90) 



dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_NKG7_vs_TIA1_SCT.pdf"))
dev.off()






```



### Subsetted vlnplots CST genes
```{r vlnplots_CST_genes}


######################
# Violin Plots
######################

output.dir <- "results/figures/Stacked_VlnPlots_CST_genes/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


######################
# for imputed genes 
######################

# set assay 
DefaultAssay(pbmc.seurat) <- "alra"

goi <- grep("^CST.", rownames(pbmc.seurat), value = T)


for(i in 1:length(cell.groups)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  temp.seurat <- subset(pbmc.seurat, idents = cell.groups[[i]])
  
  p1 <- scCustomize::Stacked_VlnPlot(temp.seurat, 
                                     features = goi,
                                     pt.size = 0, 
                                     group.by = "custom_anno",
                                     x_lab_rotate = 90) 
  
  for(j in 1:length(p1)){
    p1[[j]] <- p1[[j]] + theme(
      plot.margin = margin(t = 0,
                           r = 10,
                           b = 0,
                           l = 0, unit = "cm")
    )
  }
  
  print(p1)
  
  dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_",  names(cell.groups)[i], "group_CST_genes_imputed.pdf"))
  dev.off()
}


######################
# for rna genes 
######################

# set assay 
DefaultAssay(pbmc.seurat) <- "SCT"

goi <- grep("^CST.", rownames(pbmc.seurat), value = T)


for(i in 1:length(cell.groups)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  temp.seurat <- subset(pbmc.seurat, idents = cell.groups[[i]])
  
  p1 <- scCustomize::Stacked_VlnPlot(temp.seurat, 
                                     features = goi,
                                     pt.size = 0, 
                                     group.by = "custom_anno",
                                     x_lab_rotate = 90) 
  
  for(j in 1:length(p1)){
    p1[[j]] <- p1[[j]] + theme(
      plot.margin = margin(t = 0,
                           r = 10,
                           b = 0,
                           l = 0, unit = "cm")
    )
  }
  
  print(p1)
  
  dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_",  names(cell.groups)[i], "group_CST_genes.pdf"))
  dev.off()
}


```

### Subsetted vlnplots CTS genes
```{r vlnplots_CTS_genes}

######################
# Violin Plots
######################

output.dir <- "results/figures/Stacked_VlnPlots_CTS_genes/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}



######################
# for imputed genes 
######################

# set assay 
DefaultAssay(pbmc.seurat) <- "alra"

goi <- grep("^CTS.", rownames(pbmc.seurat), value = T)


for(i in 1:length(cell.groups)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  temp.seurat <- subset(pbmc.seurat, idents = cell.groups[[i]])
  
  p1 <- scCustomize::Stacked_VlnPlot(temp.seurat, 
                                     features = goi,
                                     pt.size = 0, 
                                     group.by = "custom_anno",
                                     x_lab_rotate = 90) 
  
  for(j in 1:length(p1)){
    p1[[j]] <- p1[[j]] + theme(
      plot.margin = margin(t = 0,
                           r = 10,
                           b = 0,
                           l = 0, unit = "cm")
    )
  }
  
  print(p1)
  
  dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_",  names(cell.groups)[i], "group_CTS_genes_imputed.pdf"))
  dev.off()
}

######################
# for rna genes 
######################

# set assay 
DefaultAssay(pbmc.seurat) <- "SCT"

goi <- grep("^CTS.", rownames(pbmc.seurat), value = T)


for(i in 1:length(cell.groups)){
  Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
  temp.seurat <- subset(pbmc.seurat, idents = cell.groups[[i]])
  
  p1 <- scCustomize::Stacked_VlnPlot(temp.seurat, 
                                     features = goi,
                                     pt.size = 0, 
                                     group.by = "custom_anno",
                                     x_lab_rotate = 90) 
  
  for(j in 1:length(p1)){
    p1[[j]] <- p1[[j]] + theme(
      plot.margin = margin(t = 0,
                           r = 10,
                           b = 0,
                           l = 0, unit = "cm")
    )
  }
  
  print(p1)
  
  dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplots_",  names(cell.groups)[i], "group_CTS_genes.pdf"))
  dev.off()
}
```

### Heatmap CST, CTS, and Cyto genes
```{r Heatmap_CST_CTS_cyto_genes}

######################
# Heatmap plot
######################

output.dir <- "results/figures/Heatmap_CST_CTS_cytogenes/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

########################
# For imputed genes
########################

# set assay 
DefaultAssay(pbmc.seurat) <- "alra"

Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
temp.seurat <- AverageExpression(pbmc.seurat, return.seurat = TRUE)


goi.1 <- grep("^CTS.", rownames(temp.seurat), value = T)
goi.2 <- grep("^CST.", rownames(temp.seurat), value = T)
goi <- c(cytotoxicity.markers, goi.1, goi.2)



DoHeatmap(temp.seurat, 
          features = goi,
          draw.lines = FALSE, 
          raster = FALSE) + 
  scico::scale_fill_scico(palette = "batlow", 
                          direction = 1, 
                          na.value = "white")

dev.copy(pdf, paste0(output.dir, "Heatmap_CST_CTS_cytogenes_imputed.pdf"))
dev.off()

########################
# For RNA genes
########################

# set assay 
DefaultAssay(pbmc.seurat) <- "SCT"

Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
temp.seurat <- AverageExpression(pbmc.seurat, return.seurat = TRUE)

goi.1 <- grep("^CTS.", rownames(temp.seurat), value = T)
goi.2 <- grep("^CST.", rownames(temp.seurat), value = T)
goi <- c(cytotoxicity.markers, goi.1, goi.2)


DoHeatmap(temp.seurat, 
          features = goi,
          draw.lines = FALSE, 
          raster = FALSE) + 
  scico::scale_fill_scico(palette = "batlow", 
                          direction = 1, 
                          na.value = "white")

dev.copy(pdf, paste0(output.dir, "Heatmap_CST_CTS_cytogenes.pdf"))
dev.off()

```

## Cytotoxicity module score & Correlation analysis {.tabset}

### Module score for cytotoxicity & Density plot
```{r cytotoxicity_and_density_plot}

DefaultAssay(pbmc.seurat) <- "SCT"

pbmc.seurat <- AddModuleScore(pbmc.seurat,
                              features = list(cytotoxicity.markers),
                              name = "Cytotoxicity_all")


######################
# plot density plots
######################

output.dir <- "results/figures/Fig_1/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

scCustomize::Plot_Density_Custom(pbmc.seurat,
                                 features = "Cytotoxicity_all1",
                                 custom_palette = rev(batlow.pal),
                                 reduction = "wnn.umap", 
                                 pt.size = 1) + 
  NoLegend()

dev.copy(pdf, paste0(output.dir, "wnnUMAP_DensityPlot_cytotoxicity_allmarkers.pdf"))
dev.off()

VlnPlot_scCustom(pbmc.seurat,
                 features = "Cytotoxicity_all1",
                 pt.size = 0,
                 group.by = "custom_anno",
                 sort = "increasing",
                 log = F) + NoLegend()

dev.copy(pdf, paste0(output.dir, "VlnPlot_cytotoxicity_allmarkers_ordered.pdf"))
dev.off()

```


### Correlation scatters
```{r correlation_scatter}

output.dir <- "results/figures/Correlation_scatter/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


DefaultAssay(pbmc.seurat) <- "SCT"
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno

avg.seurat <- AverageExpression(pbmc.seurat, 
                                group.by = "custom_anno",
                                return.seurat = TRUE)

# really unsure why the annotation gets messed up. but use below to reset the annotation to custom_anno grouping
avg.seurat@meta.data$orig.ident <- rownames(avg.seurat@meta.data)
Idents(avg.seurat) <- avg.seurat@meta.data$orig.ident



goi <- list()

for(i in seq_along(cytotoxicity.markers)){
  goi[[i]] <- cytotoxicity.markers[-i]
  name.var <- paste0("cytotoxicity_minus_", cytotoxicity.markers[i])
  names(goi)[i] <- name.var
}

goi[[length(goi)+1]] <- cytotoxicity.markers
names(goi)[length(goi)] <- "Full_cytotoxicity_sig"



avg.seurat <- AddModuleScore(avg.seurat,
                             features = goi, 
                             name = names(goi))


end.point <- length(colnames(avg.seurat@meta.data))
start.point <- end.point-length(goi)
start.point <- start.point + 1
colnames(avg.seurat@meta.data)[start.point:end.point]


library(stringr)
string.var <- colnames(avg.seurat@meta.data)[start.point:end.point]
new_string <- str_remove(string.var, ".$")
colnames(avg.seurat@meta.data)[start.point:end.point] <- new_string


for(i in seq_along(goi)){
  if(names(goi)[i] == "Full_cytotoxicity_sig"){next}
  
  feature.1.var <- names(goi)[i]
  feature.2.var <- gsub("cytotoxicity_minus_", "", feature.1.var)
  
  print(
    FeatureScatter(avg.seurat, 
                   feature1 = feature.1.var,
                   feature2 = feature.2.var,
                   pt.size = 5,
                   span = TRUE) + scale_x_continuous(limits = c(-1, 1.5))
  )
  
  dev.copy(pdf, paste0(output.dir, "Cor_scatter_", feature.1.var, "_vs_", feature.2.var, ".pdf"))
  dev.off()
}



# cor score without pDC for GZMB sig 
Idents(avg.seurat)
temp.seurat <- subset(avg.seurat, idents = "pDC", invert = T)

FeatureScatter(temp.seurat, 
               feature1 = "cytotoxicity_minus_GZMB",
               feature2 = "GZMB",
               pt.size = 5,
               span = TRUE) + scale_x_continuous(limits = c(-1, 1.5))

dev.copy(pdf, paste0(output.dir, "Cor_scatter_Cytotoxicity_minus_GZMB_vs_GZMB_without_pDC_subset.pdf"))
dev.off()



```


### Correlation Granzymes vs Granzymes
```{r Correlation_granzymes_vs_granzymes}

# Create results directories
output.dir <- "results/figures/Correlation_Granzyme_vs_Granzyme/"

if(!dir.exists(output.dir)){
  dir.create(output.dir,
             recursive = TRUE)}

# load package
require(corrplot)

##########################
# Calculate correlation
##########################


# Function
cor.mtest <- function(mat, ...) {
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}


########################
# Normalised RNA values
########################

DefaultAssay(pbmc.seurat) <- "SCT"
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$celltype.l1
table(pbmc.seurat@meta.data$custom_anno, pbmc.seurat@meta.data$celltype.l1)

# Perform analysis as all cytotoxic cells (i.e CD4/CD8/NK), then CD4, CD8, or NK individual correlation plots
group.list <- list(cytotoxic_cells = c("CD8_T", "NK"), 
                   CD8_cells = c("CD8_T"), 
                   NK_cells = c("NK"))


for(i in 1:length(group.list)){
  
  # subset dataset for particular cell types
  group.name <- names(group.list)[i]
  temp.seurat <- subset(pbmc.seurat, idents = group.list[[i]])
  
  
  # extract RNA info
  RNA.data <- temp.seurat@assays$SCT@data
  
  # reduce dataset to just cytotoxic genes
  index.vec <- grep(paste0("^", cytotoxicity.markers, "$", collapse = "|"), rownames(RNA.data))
  
  RNA.data <- RNA.data[index.vec, ]
  RNA.data <- t(as.matrix(RNA.data))
  head(RNA.data)
  rownames(RNA.data)
  
  # Calculate correlation
  corr.val <- cor(RNA.data)
  
  # Plot correlation
  
  print(
    corrplot(corr.val,
             method = "color", 
             type = "full", 
             order = "hclust", 
             tl.col = "black", 
             diag = F))
  
  dev.copy(pdf,
           paste0(output.dir, "Corplot_RNA_", group.name, ".pdf"))
  dev.off()
  
  
  # matrix of the p-value of the correlation
  p.mat <- cor.mtest(RNA.data)
  
  print(
    corrplot(corr.val,
             method = "color", 
             type = "full", 
             order = "hclust", 
             tl.col = "black", 
             diag = FALSE,
             p.mat = p.mat,
             sig.level = 0.01, 
             insig = "blank")
  )
  
  dev.copy(pdf,
           paste0(output.dir, "Corplot_RNA_", group.name, "_sig_values_only.pdf"))
  dev.off()
  
}


```

### Correlation CTS and CST genes vs Granzymes
```{r Correlation_CTS_CST_vs_Granzymes}

# Create results directories
output.dir <- "results/figures/Correlation_CTS_CST_vs_Granzymes/"

if(!dir.exists(output.dir)){
  dir.create(output.dir,
             recursive = TRUE)}

# load package
require(corrplot)

##########################
# Calculate correlation
##########################


# Function
cor.mtest <- function(mat, ...) {
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}


########################
# Normalised RNA values
########################

DefaultAssay(pbmc.seurat) <- "SCT"
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$celltype.l1
table(pbmc.seurat@meta.data$custom_anno, pbmc.seurat@meta.data$celltype.l1)

# Perform analysis as all cytotoxic cells (i.e CD8/NK), then CD4, CD8, or NK individual correlation plots
group.list <- list(cytotoxic_cells = c("CD8_T", "NK"), 
                   CD8_cells = c("CD8_T"), 
                   NK_cells = c("NK"))


##################################################
# Analysis for CTS genes vs granzyme genes 
##################################################

cts.genes <- grep("^CTS.", rownames(pbmc.seurat), value = T)
goi <- c(cytotoxicity.markers, cts.genes)


for(i in 1:length(group.list)){
  
  # subset dataset for particular cell types
  group.name <- names(group.list)[i]
  temp.seurat <- subset(pbmc.seurat, idents = group.list[[i]])
  
  
  # extract RNA info
  RNA.data <- temp.seurat@assays$SCT@data
  
  # reduce dataset to just cytotoxic genes
  index.vec <- grep(paste0("^", goi, "$", collapse = "|"), rownames(RNA.data))
  
  RNA.data <- RNA.data[index.vec, ]
  RNA.data <- t(as.matrix(RNA.data))
  head(RNA.data)
  rownames(RNA.data)
  
  # Calculate correlation
  corr.val <- cor(RNA.data)
  
  # Plot correlation
  
  print(
    corrplot(corr.val,
             method = "color", 
             type = "full", 
             order = "hclust", 
             tl.col = "black", 
             diag = F))
  
  dev.copy(pdf,
           paste0(output.dir, "Corplot_RNA_", group.name, "_CTS_vs_Granzyme.pdf"))
  dev.off()
  
  
  # matrix of the p-value of the correlation
  p.mat <- cor.mtest(RNA.data)
  
  print(
    corrplot(corr.val,
             method = "color", 
             type = "full", 
             order = "hclust", 
             tl.col = "black", 
             diag = FALSE,
             p.mat = p.mat,
             sig.level = 0.01, 
             insig = "blank")
  )
  
  
  dev.copy(pdf,
           paste0(output.dir, "Corplot_RNA_", group.name, "_CTS_vs_Granzyme_sig_values_only.pdf"))
  dev.off()
  
  
}



##################################################
# Analysis for CST genes vs granzyme genes 
##################################################
cst.genes <- grep("^CST.", rownames(pbmc.seurat), value = T)
goi <- c(cytotoxicity.markers, cst.genes)

for(i in 1:length(group.list)){
  
  # subset dataset for particular cell types
  group.name <- names(group.list)[i]
  temp.seurat <- subset(pbmc.seurat, idents = group.list[[i]])
  
  
  # extract RNA info
  RNA.data <- temp.seurat@assays$SCT@data
  
  # reduce dataset to just cytotoxic genes
  index.vec <- grep(paste0("^", goi, "$", collapse = "|"), rownames(RNA.data))
  
  RNA.data <- RNA.data[index.vec, ]
  RNA.data <- t(as.matrix(RNA.data))
  head(RNA.data)
  rownames(RNA.data)
  
  # Calculate correlation
  corr.val <- cor(RNA.data)
  
  # Plot correlation
  
  print(
    corrplot(corr.val,
             method = "color", 
             type = "full", 
             order = "hclust", 
             tl.col = "black", 
             diag = F))
  
  dev.copy(pdf,
           paste0(output.dir, "Corplot_RNA_", group.name, "_CST_vs_Granzyme.pdf"))
  dev.off()
  
  
  # matrix of the p-value of the correlation
  p.mat <- cor.mtest(RNA.data)
  
  print(
    corrplot(corr.val,
             method = "color", 
             type = "full", 
             order = "hclust", 
             tl.col = "black", 
             diag = FALSE,
             p.mat = p.mat,
             sig.level = 0.01, 
             insig = "blank")
  )
  
  
  dev.copy(pdf,
           paste0(output.dir, "Corplot_RNA_", group.name, "_CST_vs_Granzyme_sig_values_only.pdf"))
  dev.off()
  
  
}


```
 
 
 
## Co-expression boolean {.tabset}



### Co-expression boolean upset plots
```{r boolean_coexpression_upset_plots}

# note use ComplexUpset not UpsetR package
#install.packages('ComplexUpset')
library(ComplexUpset)

# Create results directories
output.dir <- "results/figures/PBMC_dataset_boolean_upset_plots/"

if(!dir.exists(output.dir)){
  dir.create(output.dir,
             recursive = TRUE)}

################################################
# Boolean expression patterns within cell types
################################################

# determine thresholds for expression 
DefaultAssay(pbmc.seurat) <- "SCT"

theshold.val <- 1

for(i in seq_along(cytotoxicity.markers)){
  print(
  scCustomize::VlnPlot_scCustom(pbmc.seurat, 
                              feature = cytotoxicity.markers[i], 
                              pt.size = 0, 
                              group.by = "custom_anno") + 
  NoLegend() +
  geom_hline(yintercept = theshold.val, 
             linetype = "dashed",
             color = "red")
  )
  dev.copy(pdf, paste0(output.dir, "Threshold_val_determination_", cytotoxicity.markers[i], ".pdf"))
  dev.off()
  
}



########################################
# perform analysis for CD8 T cells
########################################

# CD4 CTL
# CD8 EM
# NK Bright
# NK Dim
#gd 
#gdVD2
#MAIT


# Determine expression status
DefaultAssay(pbmc.seurat) <- "SCT"
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
unique(pbmc.seurat@meta.data$custom_anno)


# populations to loop over
subset.var <- c("CD8_Naive", "CD8_CM", "dnT", "CD8_EM", "CD4_CTL", "NK_Bright", "NK_Dim", "gdT_V9D2", "gdT", "MAIT")

for(j in seq_along(subset.var)){
  
  print(paste0("Performing analysis for ", subset.var[j]))
  
  # subset dataset 
  temp.seurat <- subset(pbmc.seurat, idents = subset.var[j])
  
  # quantify granzyme expression
  for(i in seq_along(cytotoxicity.markers)){
    logic.vec <- rownames(temp.seurat@assays$SCT@data) == cytotoxicity.markers[i]
    goi.expression <- temp.seurat@assays$SCT@data[logic.vec, ]
    expression.logic <- goi.expression > theshold.val
    sum(names(expression.logic) != rownames(temp.seurat@meta.data)) # same order, nice and easy 
    
    temp.seurat@meta.data[[paste0(cytotoxicity.markers[i], "_status")]] <- expression.logic
    
  }
  
  # extract and format metadata
  meta.data <- temp.seurat@meta.data %>%
    dplyr::select(ends_with("_status"))
  
  input_data <- data.frame(lapply(meta.data, as.integer))
  
  rownames(input_data) <- rownames(meta.data)
  
  colnames(input_data) <- gsub("_status", "", colnames(input_data))
  
  print(
    # plot upset plot 
    ComplexUpset::upset(data = input_data, 
                        n_intersections = 10,
                        intersect = rev(c("GZMA", "GZMB", "GZMH", "GZMK", "GZMM", "GNLY", "PRF1", "NKG7")),
                        name = "cytotoxic profile",
                        width_ratio = 0.15,
                        height_ratio = 0.5,
                        min_size = 1,
                        keep_empty_groups = TRUE,
                        sort_intersections = 'descending',
                        sort_sets = FALSE,
                        base_annotations = list('Intersection size' = intersection_size(
                          text_mapping = aes(label = paste0(round((!!get_size_mode("exclusive_intersection")/nrow(input_data))*100, 2), "%",
                                                            "\n",
                                                            "(",
                                                            !!get_size_mode("exclusive_intersection"),
                                                            ")")))
                          
                          + annotate(
                            geom = 'text',
                            x = Inf, 
                            y = Inf,
                            label = paste('Total:', nrow(input_data)),
                            vjust = 1,
                            hjust = 1)
                          
                          + ylab('Intersection size')
                        ), 
                        set_sizes = (upset_set_size()
                                     + theme(axis.ticks.x = element_line())
                                     + scale_y_continuous(trans = reverse_log_trans(),
                                                          labels = log10)
                                     + ylab("log10(set size)"))
    )
    
  )
  
  
  dev.copy(pdf, paste0(output.dir, "Upset_plot_", subset.var[j], "_top10_intersects.pdf"))
  dev.off()
  
  
}





```
 
 
### Co-expression boolean heatmap plots

```{r boolean_heatmap_plots}



# Create results directories
output.dir <- "results/figures/PBMC_dataset_boolean_heatmaps/"

if(!dir.exists(output.dir)){
  dir.create(output.dir,
             recursive = TRUE)}




library(gplots)

######################################################################
# heatmaps of non-scaled values showing co-expression patterns 
######################################################################

# set up object
DefaultAssay(pbmc.seurat) <- "SCT"
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno


# populations to loop over
subset.var <- c("CD8_Naive", "CD8_CM", "dnT", "CD8_EM", "CD4_CTL", "NK_Bright", "NK_Dim", "gdT_V9D2", "gdT", "MAIT")

table(pbmc.seurat@meta.data$custom_anno)

# set a custom breaks so all heatmaps are on same colour scale and can be compared
length(batlow.pal)

custom_breaks <- seq(from = 0, to = 6, length.out = 101)


for(i in seq_along(subset.var)){
  
  # subset dataset
  temp.seurat <- subset(pbmc.seurat, idents = subset.var[i])
  
  # scale dataset
  temp.seurat <- ScaleData(temp.seurat)
  
  
  # extract values 
  data_values <- FetchData(temp.seurat,
                           vars = cytotoxicity.markers, 
                           layer = "data")
  
  # remove any NA values
  data_values <- data_values %>%
    filter(!all(is.na(.)))  # Keep rows where not all values are NA
  
  
  # set random sample value, min of 2,000 or entire dataset
  n.val <- min(2000, nrow(data_values))  
  
  
  # Randomly sample rows
  sampled_df <- data_values %>% 
    sample_n(size = n.val, replace = FALSE)  # Sample without replacement
  
  # flip dataframe
  sampled_df <- t(sampled_df)
  
  # if value below threshold value, set to 0
  sampled_df[sampled_df < theshold.val] <- 0
  
  
  # Create heatmap
  print(
    heatmap.2(as.matrix(sampled_df),   # Convert data frame to matrix
              trace = "none",   # Do not display trace lines
              col = batlow.pal,  
              breaks = custom_breaks,
              scale = "none",   # Do not scale rows or columns
              dendrogram = "column",  
              Rowv = FALSE,     
              Colv = TRUE,     
              margins = c(10, 10),  # Adjust margins for row and column labels
              labRow = rownames(sampled_df),
              labCol = TRUE,
              na.rm=TRUE,
              na.color = "white",
              xlab = "Cells",
              ylab = "Cytotoxicity Markers",
              main = paste0(subset.var[i], " not scaled"),
              distfun = function(x){dist(x , method = "euclidean")},
              hclustfun = function(x){hclust(x , method = "ward.D2")}
    )
  )
  
  dev.copy(pdf, paste0(output.dir, "Heatmap_not_scaled_", subset.var[i], "_only.pdf"))
  dev.off()
  
  
}




```
 
 
 
 ### TEMP KEEP< BUT PROBABLY REMOVE this chunk
 
### Visualise NKG7 and TIA1 expression in bulk RNAseq datasets
```{r bulk_RNAseq_datasets_NKG7_vs_TIA1}


library(ggplot2)


Monaco_data <- celldex::MonacoImmuneData()



# Extract gene expression data
gene_expression <- assay(Monaco_data)
gene_expression <- as.data.frame(gene_expression)

gene_expression$gene <- rownames(gene_expression)


grep("TIA1", rownames(gene_expression), value = T)
grep("NKG7", rownames(gene_expression), value = T)

gene_expression <- gene_expression[rownames(gene_expression) %in% c("NKG7", "TIA1"), ]


library(tidyr)
library(ggplot2)

# Convert data to long format
df_long <- pivot_longer(gene_expression, 
                        cols = -gene, 
                        names_to = "Cell_Type",
                        values_to = "Expression")


unique(df_long$Cell_Type)



# Remove alphanumeric code and first underscore from Cell_Type
df_long$Cell_Type <- gsub("^[[:alnum:]]{4}_", "", df_long$Cell_Type)

# Display unique cell types after modification
unique(df_long$Cell_Type)



# Calculate average expression value for NKG7 for each cell type
average_expression <- df_long %>%
  filter(gene == "NKG7") %>%
  group_by(Cell_Type) %>%
  summarise(Avg_Expression_NKG7 = mean(Expression))

# Reorder levels of Cell_Type factor based on average expression value for NKG7
df_long$Cell_Type <- factor(df_long$Cell_Type, levels = average_expression$Cell_Type[order(average_expression$Avg_Expression_NKG7, decreasing = T)])

# Plot boxplot
ggplot(df_long, aes(x = Cell_Type, y = Expression, fill = gene)) +
  geom_boxplot() +
  geom_jitter(position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "Cell Type", y = "Expression Value", fill = "Gene") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("NKG7" = "red", "TIA1" = "blue"))  # Customize colors if needed



```





 ## needs correcting 
### NKG7 correlation across all genes 
```{r correlation_of_NKG7_vs_all_genes}


# Create results directories
output.dir <- "results/figures/Correlation_goi_vs_all_genes/"

if(!dir.exists(output.dir)){
  dir.create(output.dir,
             recursive = TRUE)}


########################
# Normalised RNA values
########################

DefaultAssay(pbmc.seurat) <- "SCT"
Idents(pbmc.seurat) <- pbmc.seurat@meta.data$celltype.l1
table(pbmc.seurat@meta.data$custom_anno, pbmc.seurat@meta.data$celltype.l1)

table(pbmc.seurat@meta.data$celltype.l1)
# 25K CD8, 18K NK cells 


group.list <- list(cytotoxic_cells = c("CD8_T", "NK"), 
                   CD8_cells = c("CD8_T"), 
                   NK_cells = c("NK"))
i = 3
output.list <- list()

  for(i in length(group.list)){
  # subset dataset for particular cell types
  group.name <- names(group.list)[i]
  temp.seurat <- subset(pbmc.seurat, idents = group.list[[i]])
  
  
  # extract normalised data table
normed.data <- as.matrix(temp.seurat@assays$SCT@data)

# calculate correlation
output.list[[i]] <- sc.correlation(data.slot = normed.data,
                        goi = "NKG7",
                        remove.quantile = 0.3,
                        output.tables = paste0("results/tables/Correlation_analysis_", group.name, "/"))

names(output.list)[i] <- group.name

  }



top.genes <- output.list[[3]] %>%
  dplyr::filter(FDR < 0.05) %>%
  dplyr::top_n(50, Cor.val) %>%
  dplyr::arrange(desc(Cor.val)) %>%
  pull(GeneID)


bot.genes <- output.list[[3]] %>%
  dplyr::filter(FDR < 0.05) %>%
  dplyr::top_n(-50, Cor.val) %>%
  dplyr::arrange(Cor.val) %>%
  pull(GeneID)

goi <- c(top.genes, bot.genes)







temp.seurat <- subset(temp.seurat, downsample = 500)
temp.seurat <- ScaleData(temp.seurat)

DoHeatmap(temp.seurat, 
          group.by = "custom_anno", 
          features = goi)




scCustomize::VlnPlot_scCustom(temp.seurat, "CST7", pt.size = 0)

goi.1 <- grep("^CST.", rownames(temp.seurat), value = T)
goi.2 <- grep("^CTS.", rownames(temp.seurat), value = T)

Idents(pbmc.seurat) <- pbmc.seurat@meta.data$custom_anno
temp.seurat <- AverageExpression(pbmc.seurat, return.seurat = T)


DoHeatmap(temp.seurat, 
          draw.lines = F,
          features = c(goi.1, goi.2, cytotoxicity.markers), raster = F)


# Getting the frequency count using table
freq_table <- table(pbmc.seurat@meta.data$custom_anno)

# Converting to data frame
df <- data.frame(value = names(freq_table), frequency = as.numeric(freq_table))

# Reshaping to long format
df_long <- df %>%
  arrange(desc(frequency))  # Arranging by frequency in descending order

# Show the transformed data


df_long$value <- factor(df_long$value, levels = df_long$value)



ggplot(df_long, aes(x = value, y = frequency)) +
  geom_col(fill = "#112446") +
  theme_minimal() +
  #scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



corrplot(corr.val,
         method = "shade", 
         type = "full", 
         order = "hclust", 
         tl.col = "black", 
         diag = FALSE,
         p.mat = p.mat,
         sig.level = 0.01, 
         insig = "blank")


DefaultAssay(temp.seurat) <- "alra"
FeatureScatter(temp.seurat, 
               feature1 = "GZMK",
               feature2 = "GZMB",
               pt.size = 5,
               span = TRUE)








```



### WGCNA
```{r remedy001}

# run in terminal 
# create new conda environment for R
#conda create -n hdWGCNA -c conda-forge r-base r-essentials

# activate conda environment
#conda activate hdWGCNA

```




### analysed up to here


### Subsettting dataset for NK, CD8, and CD4 subsets and perform dim reduction 
```{r subsetting_dataset}


# Set up dataset
DefaultAssay(reference) <- "SCT"
Idents(reference) <- reference@meta.data$celltype.l1
unique(Idents(reference))


# Subset dataset
#CD8.seurat <- subset(reference, idents = "CD8 T")
#CD4.seurat <- subset(reference, idents = "CD4 T")
#NK.seurat <- subset(reference, idents = "NK")

# Make a list of seurat objects
#seurat.list <- c(CD8.seurat, CD4.seurat, NK.seurat)
#names(seurat.list) <- c("CD8_cells", "CD4_cells", "NK_cells")

NK.T.seurat <- subset(reference, idents = c("CD8 T", "CD4 T", "NK"))

seurat.list <- c(NK.T.seurat)
names(seurat.list) <- c("NK_T_cells")

rm(NK.T.seurat)


# remove single seurat objects
#rm(CD8.seurat, CD4.seurat, NK.seurat)




# Re-normalise
seurat.list <- lapply(X = seurat.list, 
                      FUN = SCTransform, 
                      assay = "SCT",
                      method = "glmGamPoi", 
                      verbose = TRUE) 


# Scale and Calculate PCA - RNA
for(i in seq_along(seurat.list)){
  seurat.list[[i]] <- seurat.list[[i]] %>%
    NormalizeData() %>%
    FindVariableFeatures() %>% 
    ScaleData() %>% 
    RunPCA()
}






# Normalise, Scale and Calculate PCA - ADT

for(i in seq_along(seurat.list)){
  
  DefaultAssay(seurat.list[[i]]) <- 'ADT'
  
  # we will use all ADT features for dimensional reduction
  # we set a dimensional reduction name to avoid overwriting the 
  VariableFeatures(seurat.list[[i]]) <- rownames(seurat.list[[i]][["ADT"]])
  
  seurat.list[[i]] <- seurat.list[[i]] %>%
    NormalizeData(normalization.method = 'CLR', 
                  margin = 2) %>% 
    ScaleData() %>% 
    RunPCA(reduction.name = 'apca')
}


seurat.list <- lapply(X = seurat.list, 
                      FUN = FindMultiModalNeighbors, 
                      reduction.list = list("pca", "apca"),
                      dims.list = list(1:30, 1:18), 
                      modality.weight.name = "Multi.weight") 

seurat.list <- lapply(X = seurat.list, 
                      FUN = RunSPCA, 
                      assay = 'SCT',
                      graph = 'wsnn') 


seurat.list <- lapply(X = seurat.list, 
                      FUN = RunUMAP, 
                      reduction = 'spca', 
                      dims = 1:50) 




DimPlot(seurat.list[[1]], group.by = "celltype.l1", reduction = "umap") 

DimPlot(seurat.list[[1]], group.by = "donor", reduction = "umap") 
DimPlot(seurat.list[[1]], group.by = "donor", reduction = "wnn.umap") 
DimPlot(seurat.list[[1]], group.by = "celltype.l2", reduction = "umap") 



DimPlot(object = NK.seurat, 
        reduction = "wnn.umap",
        group.by = "celltype.l1",
        label = TRUE, 
        label.size = 3, 
        repel = TRUE) + 
  NoLegend()









DimPlot(reference, group.by = "donor", reduction = "umap") 
DimPlot(reference, group.by = "celltype.l2", reduction = "umap") 




SaveH5Seurat(temp.seurat, 
             filename = "Exported_RDS_files/pbmc_multimodal_CD8_cells_only.h5seurat",
             overwrite = TRUE)

# remove data
rm(temp.seurat)



```

### work in progress, playground 

### heatmap
```{r heatmap_TIA_NKG7}

######################
# Heatmap plot
######################

output.dir <- "results/figures/NKG7_vs_TIA1/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

########################
# For RNA genes
########################

# set assay 
DefaultAssay(pbmc.seurat) <- "SCT"

goi <- c("TIA1", "NKG7")

# quick plot to look at data
VlnPlot_scCustom(pbmc.seurat,
                 features = goi,
                 pt.size = 0,
                 group.by = "celltype.l1",
                 sort = "increasing",
                 log = FALSE, 
                 same.y.lim = T) + NoLegend()

# select order for data to be in 
temp.seurat <- pbmc.seurat
temp.seurat@meta.data$celltype.l1 <- factor(temp.seurat@meta.data$celltype.l1, levels = c("NK", "other_T", "CD8_T", "other", "CD4_T", "B", "Mono", "DC"))

# ordered data
VlnPlot_scCustom(temp.seurat,
                 features = goi,
                 pt.size = 0,
                 group.by = "celltype.l1",
                 log = FALSE, 
                 same.y.lim = T) + NoLegend()

dev.copy(pdf, paste0(output.dir, "VlnPlot_NKG7_vs_TIA1.pdf"))
dev.off()



# data from gene expression atlas



```


