---
title: "Tabula_Sapeins_dataset_analysis"
author: "Dillon Corvino"
date: "04/02/2024"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    highlight: tango
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Built with R version `r getRversion()`

## Setup {.tabset}

### Dataset information
```{r Dataset_Info}



```

### Environment
```{r setup}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,  
  eval = TRUE, 
  tidy = TRUE
)

knitr::opts_knit$set(
  root.dir = "../"
)


# Environment Set up
rm(list = ls()) #Clean workspace
cat("\014")     #Clean Console
gc() # Free memory

# pipeline variables
long.compute <- FALSE

# Establish working directory
library("rstudioapi")

# Set working directory to source file location
setwd(dirname(getActiveDocumentContext()$path))
setwd("..")

# Create output directories & load custom functions & colour scheme
source("scripts/Setup.R", local = knitr::knit_global())

```


### load dataset
```{r load_dataset}

#######################
# load dataset
#######################

# dataset processed as part of Tabula_Sapiens github repo, filter and normalise and integrate and caclulate umap dims etc all done in that script

ts.immune.seurat <- LoadH5Seurat("data/TS_immune_imputed.h5Seurat", assays = c("SCT", "alra"))

################################################################################################
# for easy of use rename the integrattted cca umap in the object to just "umap"
################################################################################################

new_reduction <- CreateDimReducObject(embeddings = Embeddings(ts.immune.seurat, reduction = "umap.cca.sct"), key = "umap")
ts.immune.seurat[["umap"]] <- new_reduction

```

## Visualise {.tabset}

### Featureplots, density plots and vlnplots across whole dataset
```{r vis_cytotoxic_across_whole_dataset}

DefaultAssay(ts.immune.seurat) <- "alra"

cts.genes <- grep("CTS.*", rownames(ts.immune.seurat), value = T)
goi <- c(cytotoxicity.markers, cts.genes)

######################
# plot featureplots
######################

output.dir <- "results/figures/tabula_sapiens_FeaturePlots/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


for(i in seq_along(goi)){
  print(
    FeaturePlot(ts.immune.seurat,
                features = goi[i],
                order = TRUE, 
                pt.size = 1, 
                raster = TRUE) + 
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "UMAP_FeaturePlot_", goi[i], ".pdf"))
  dev.off()
  
}


######################
# plot density plots
######################

output.dir <- "results/figures/tabula_sapiens_DensityPlots/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

for(i in seq_along(goi)){
  print(
    scCustomize::Plot_Density_Custom(ts.immune.seurat,
                                     features = goi[i],
                                     custom_palette = batlow.pal,
                                     pt.size = 1) + 
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "UMAP_DensityPlot_", goi[i], ".pdf"))
  dev.off()
  
}


######################
# Violin Plots
######################
output.dir <- "results/figures/tabula_sapiens_Stacked_VlnPlots/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

head(ts.immune.seurat@meta.data)

vars.plot <- c("organ_tissue", "gender", "Grouped_l1", "Grouped_l2", "Grouped_l3", "Grouped_l4")

for(i in seq_along(vars.plot)){
  print(
    scCustomize::Stacked_VlnPlot(ts.immune.seurat, 
                                 features = cytotoxicity.markers,
                                 pt.size = 0, 
                                 group.by = vars.plot[i],
                                 raster = TRUE,
                                 x_lab_rotate = TRUE)
  )
  
  dev.copy(pdf, paste0(output.dir, "Stacked_vlnplots_cytotoxicity_markers_", vars.plot[i], ".pdf"))
  dev.off()
}





```


### CD8, CD4, and NK cyto expression across immune specific organs - heatmaps and vlnplots
```{r immune_specific_organs_cyto_expression}

#################################
# Immune specific organs 
#################################

DefaultAssay(ts.immune.seurat) <- "alra"

output.dir <- "results/figures/tabula_sapiens_immune_specific_organs/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


# identify immune organs
immune.organs <- c("Spleen", "Thymus", "Salivary_Gland", "Lymph_Node", "Blood", "Bone_Marrow")

# set idents
Idents(ts.immune.seurat) <- ts.immune.seurat@meta.data$organ_tissue

# subset by immune organs
immune.organ.seurat <- subset(ts.immune.seurat, idents = immune.organs)
immune.organ.seurat # 173,153 cells

table(immune.organ.seurat@meta.data$gender, immune.organ.seurat@meta.data$organ_tissue)
table(immune.organ.seurat@meta.data$Grouped_l3, immune.organ.seurat@meta.data$organ_tissue)
table(immune.organ.seurat@meta.data$Grouped_l2, immune.organ.seurat@meta.data$organ_tissue)
table(immune.organ.seurat@meta.data$Grouped_l4, immune.organ.seurat@meta.data$organ_tissue)

# at level 3 resolution 
# iteratively investigate these cells across tissues
cell.types <- c("NK", "CD8_Tcell", "CD4_Tcell")

# set idents
Idents(immune.organ.seurat) <- immune.organ.seurat@meta.data$Grouped_l3

for(i in seq_along(cell.types)){
  
temp.seurat <- subset(immune.organ.seurat, idents = cell.types[i])

pseudo.ts <- AggregateExpression(temp.seurat, 
                                 assays = "SCT",
                                 return.seurat = TRUE,
                                 group.by = c("organ_tissue"))

pseudo.ts <- ScaleData(pseudo.ts)

head(pseudo.ts@meta.data)

# set idents of pseudobulk object
Idents(pseudo.ts) <- "orig.ident"


# heatmap of cytotoxic markers
print(
DoHeatmap(pseudo.ts, 
          features = cytotoxicity.markers, 
          draw.lines = FALSE, 
          raster = FALSE) + 
  NoLegend() +  
  scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")
)

dev.copy(pdf, paste0(output.dir, "Heatmap_cytotoxic_molecules_across_tissue_", cell.types[i], "_pseudobulk.pdf"))
dev.off()


DefaultAssay(temp.seurat) <- "alra"

print(
scCustomize::Stacked_VlnPlot(temp.seurat, 
                             features = cytotoxicity.markers,
                             pt.size = 0, 
                             group.by = "organ_tissue",
                             raster = FALSE,
                             split.plot = FALSE,
                             x_lab_rotate = TRUE)
)

dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplot_cytotoxic_molecules_across_tissue_", cell.types[i], ".pdf"))
dev.off()


}



```

### Gender differences
```{r gender_differences}

output.dir <- "results/figures/tabula_sapiens_gender_differences/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


# identify immune organs
immune.organs <- c("Spleen", "Thymus", "Salivary_Gland", "Lymph_Node", "Blood", "Bone_Marrow")

# set idents
Idents(ts.immune.seurat) <- ts.immune.seurat@meta.data$organ_tissue

# subset by immune organs
immune.organ.seurat <- subset(ts.immune.seurat, idents = immune.organs)
immune.organ.seurat # 173,153 cells

table(immune.organ.seurat@meta.data$gender, immune.organ.seurat@meta.data$Grouped_l3)

# specifically these cytotoxic subsets
cell.types <- c("NK", "CD8_Tcell")

# set idents
Idents(immune.organ.seurat) <- immune.organ.seurat@meta.data$Grouped_l3

temp.seurat <- subset(immune.organ.seurat, idents = cell.types)


table(temp.seurat@meta.data$gender, temp.seurat@meta.data$Grouped_l3)


DefaultAssay(temp.seurat) <- "alra"

temp.seurat@meta.data$custom_anno <- paste0(temp.seurat@meta.data$Grouped_l3, "_", temp.seurat@meta.data$organ_tissue)


scCustomize::Stacked_VlnPlot(temp.seurat, 
                             features = cytotoxicity.markers,
                             pt.size = 0, 
                             group.by = "custom_anno",
                             split.by = "gender",
                             raster = TRUE,
                             split.plot = FALSE,
                             x_lab_rotate = TRUE)

dev.copy(pdf, paste0(output.dir, "Stacked_Vlnplot_cytotoxic_molecules_across_tissue_immune_relevant_tissue_and_select_celltypes_split_gender_female_blue.pdf"))
dev.off()

# pseudobulk for heatmap

pseudo.ts <- AggregateExpression(temp.seurat, 
                                 assays = "SCT",
                                 return.seurat = TRUE,
                                 group.by = c("Grouped_l3", "organ_tissue", "gender"))


pseudo.ts <- ScaleData(pseudo.ts)

head(pseudo.ts@meta.data)

# set idents of pseudobulk object
Idents(pseudo.ts) <- "orig.ident"


# heatmap 
DoHeatmap(pseudo.ts, 
          features = cytotoxicity.markers, 
          draw.lines = FALSE, 
          raster = TRUE) + NoLegend() +  
  scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")

dev.copy(pdf, paste0(output.dir, "Heatmap_cytotoxic_molecules_across_celltype_tissue_gender_pseudobulk.pdf"))
dev.off()






```
 



## Cytotoxicity module score & Correlation analysis {.tabset}


### Correlation scatters
```{r correlation_scatter}

output.dir <- "results/figures/tabula_sapiens_Correlation_scatter/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}



DefaultAssay(ts.immune.seurat) <- "SCT"

# just NK and CD8 T cells
# set idents
Idents(ts.immune.seurat) <- ts.immune.seurat@meta.data$Grouped_l3
unique(Idents(ts.immune.seurat))

# subset by immune organs
NK.seurat <- subset(ts.immune.seurat, idents = c("NK", "CD8_Tcell"))
NK.seurat # 33,467 cells
 


pseudo.ts <- AggregateExpression(NK.seurat, 
                                 assays = "SCT",
                                 return.seurat = TRUE,
                                 group.by = c("Grouped_l3", "organ_tissue"))


Idents(pseudo.ts) <- pseudo.ts@meta.data$orig.ident
length(unique(Idents(pseudo.ts))) # 31 classifications


goi <- list()

for(i in seq_along(cytotoxicity.markers)){
  goi[[i]] <- cytotoxicity.markers[-i]
  name.var <- paste0("cytotoxicity_minus_", cytotoxicity.markers[i])
  names(goi)[i] <- name.var
}

goi[[length(goi)+1]] <- cytotoxicity.markers
names(goi)[length(goi)] <- "Full_cytotoxicity_sig"



pseudo.ts <- AddModuleScore(pseudo.ts,
                             features = goi, 
                             name = names(goi))


end.point <- length(colnames(pseudo.ts@meta.data))
start.point <- end.point-length(goi)
start.point <- start.point + 1
colnames(pseudo.ts@meta.data)[start.point:end.point]


library(stringr)
string.var <- colnames(pseudo.ts@meta.data)[start.point:end.point]
new_string <- str_remove(string.var, ".$")
colnames(pseudo.ts@meta.data)[start.point:end.point] <- new_string


for(i in seq_along(goi)){
  if(names(goi)[i] == "Full_cytotoxicity_sig"){next}
  
  feature.1.var <- names(goi)[i]
  feature.2.var <- gsub("cytotoxicity_minus_", "", feature.1.var)
  
  print(
    FeatureScatter(pseudo.ts, 
                   feature1 = feature.1.var,
                   feature2 = feature.2.var,
                   pt.size = 5,
                   span = TRUE) + 
      scale_x_continuous(limits = c(-0.5, 1.5)) +
      NoLegend()
  )
  
  dev.copy(pdf, paste0(output.dir, "Cor_scatter_", feature.1.var, "_vs_", feature.2.var, ".pdf"))
  dev.off()
}


FeatureScatter(pseudo.ts, 
               feature1 = feature.1.var,
               feature2 = feature.2.var,
               pt.size = 5,
               span = TRUE)
dev.copy(pdf, paste0(output.dir, "cor_scatter_legend.pdf"))
dev.off()

```


### Thymic T-cell development and Treg across tissues
```{r Thymic_Tcell_Development_Treg_across_tissues}

output.dir <- "results/figures/tabula_sapiens_thymic_tcell_development/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}


# subset for thymus
Idents(ts.immune.seurat) <- ts.immune.seurat@meta.data$organ_tissue
temp.seurat <- subset(ts.immune.seurat, idents = "Thymus")

# subset for thymic t-cell subsets
Idents(temp.seurat) <- temp.seurat@meta.data$Grouped_l4
table(temp.seurat@meta.data$Grouped_l4)

temp.seurat <- subset(temp.seurat, idents = c("DN1", "DN2", "DN3", "CD4_Tcell", "CD8_innate_like_thymic_resident", "CD8_Tcell", "TFH", "Thymocyte", "Treg"))

# plot 
Stacked_VlnPlot(temp.seurat, 
                cytotoxicity.markers, 
                group.by = "Grouped_l4",
                pt.size = 0, 
                assay = "alra") + NoLegend()

dev.copy(pdf, paste0(output.dir, "stacked_vln_thymic_only_Tcell_development.pdf"))
dev.off()

# subset for thymic DN t-cell subsets
Idents(temp.seurat) <- temp.seurat@meta.data$Grouped_l4
table(temp.seurat@meta.data$Grouped_l4)

temp.seurat <- subset(temp.seurat, idents = c("DN1", "DN2", "DN3"))

# plot 
Stacked_VlnPlot(temp.seurat, 
                cytotoxicity.markers, 
                group.by = "Grouped_l4",
                pt.size = 0, 
                assay = "alra") + NoLegend()

dev.copy(pdf, paste0(output.dir, "stacked_vln_thymic_only_Tcell_development_DN_subsets.pdf"))
dev.off()

####################
# Treg across tissue
####################

output.dir <- "results/figures/tabula_sapiens_Treg_across_tissue/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}

# subset for DN t-cell stages
Idents(ts.immune.seurat) <- ts.immune.seurat@meta.data$Grouped_l4

temp.seurat <- subset(ts.immune.seurat, idents = c("Treg"))

table(temp.seurat@meta.data$Grouped_l4, temp.seurat@meta.data$organ_tissue)

Stacked_VlnPlot(temp.seurat, 
                cytotoxicity.markers,
                group.by = "organ_tissue", 
                pt.size = 0,
                assay = "alra") + NoLegend()


dev.copy(pdf, paste0(output.dir, "stacked_vln_treg_expression_across_tissue.pdf"))
dev.off()



```









# not written below here
### cytotoxic molecule expression for celltypes across tissues
```{r cytotoxic_expression_across_tissues}

DefaultAssay(ts.immune.seurat) <- "alra"



output.dir <- "results/figures/across_tissues/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}




# same cell across tissue

vars.plot <- c("organ_tissue", "gender", "Grouped_l1", "Grouped_l2", "Grouped_l3", "Grouped_l4")

length(unique(ts.immune.seurat@meta.data$organ_tissue)) # 23
length(unique(ts.immune.seurat@meta.data$donor)) # 12

length(unique(ts.immune.seurat@meta.data$Grouped_l1)) # 4
length(unique(ts.immune.seurat@meta.data$Grouped_l2)) #15
length(unique(ts.immune.seurat@meta.data$Grouped_l3)) # 29
length(unique(ts.immune.seurat@meta.data$Grouped_l4)) # 67


table(ts.immune.seurat@meta.data$donor, ts.immune.seurat@meta.data$organ_tissue)
table(ts.immune.seurat@meta.data$donor, ts.immune.seurat@meta.data$Grouped_l3)
table(ts.immune.seurat@meta.data$Grouped_l3, ts.immune.seurat@meta.data$organ_tissue)

Idents(ts.immune.seurat) <- ts.immune.seurat@meta.data$Grouped_l4
unique(Idents(ts.immune.seurat))

temp.seurat <- subset(ts.immune.seurat, idents = "CD8_Tcell")
temp.seurat
table(temp.seurat@meta.data$gender, temp.seurat@meta.data$organ_tissue)

pseudo.ts <- AggregateExpression(temp.seurat, 
                                 assays = "SCT",
                                 return.seurat = TRUE,
                                 group.by = c("organ_tissue"))


pseudo.ts <- ScaleData(pseudo.ts)

head(pseudo.ts@meta.data)

# set idents of pseudobulk object
Idents(pseudo.ts) <- "orig.ident"


# heatmap of top 10 degs per cluster
DoHeatmap(pseudo.ts, 
          features = cytotoxicity.markers, 
          draw.lines = FALSE, 
          raster = TRUE) + NoLegend() +  
  scico::scale_fill_scico(palette = "batlow", 
                              direction = 1, 
                              na.value = "white")

dev.copy(pdf, paste0(output.dir, "Heatmap_cytotoxic_molecules_across_tissue_pseudobulk.pdf"))
dev.off()

DefaultAssay(temp.seurat) <- "alra"


scCustomize::Stacked_VlnPlot(temp.seurat, 
                             features = cytotoxicity.markers,
                             pt.size = 0, 
                             group.by = "organ_tissue",
                             #split.by = "gender",
                             raster = TRUE,
                             split.plot = FALSE,
                             x_lab_rotate = TRUE)





```
 






### Broad overview of dataset
```{r dataset_overview}


output.dir <- "results/data_overview/"
if(!dir.exists(output.dir)){
  dir.create(output.dir, 
             recursive = T)}



library(ggplot2)


head(ts.immune.seurat@meta.data)


if(do.plot){
  
  vars.plot <- c("donor", "organ_tissue", "gender", "anatomical_information", "Grouped_l1", "Grouped_l2", "Grouped_l3", "Grouped_l4", "seurat_clusters")
  
  for(i in seq_along(vars.plot)){
    
    print(scCustomize::DimPlot_scCustom(ts.immune.seurat, 
                                        reduction = "umap.cca.sct", 
                                        group.by = vars.plot[i], 
                                        pt.size = 1,
                                        label = TRUE,
                                        figure_plot = TRUE)
    )
    
    
    dev.copy(pdf, paste0(output.dir, "UMAP_integrated_tabula_sapiens_dataset_grouped_by_", vars.plot[i], ".pdf"))
    dev.off()
    
  }
}


################################################
# plot some distribution statistics 
################################################


# gender distribution
x <- table(ts.immune.seurat@meta.data$gender)
y <- x/sum(x)
y <- y*100

y <- as.data.frame(y)

ggplot(y) +
  aes(x = Var1, y = Freq, fill = Var1) +
  geom_col() +
  scale_fill_hue(direction = 1) +
  labs(y = "Freq of sequenced cells", x = "Gender") +
  theme_minimal() +
  ylim(0, 100)

dev.copy(pdf, paste0(output.dir, "Frequency_of_sequenced_cells_by_Gender.pdf"))
dev.off()


# Donor distribution
x <- table(ts.immune.seurat@meta.data$donor)
y <- x/sum(x)
y <- y*100

y <- as.data.frame(y)

ggplot(y) +
  aes(x = Var1, y = Freq, fill = Var1) +
  geom_col() +
  scale_fill_hue(direction = 1) +
  labs(y = "Freq of sequenced cells", x = "Donor") +
  theme_minimal() +
  ylim(0, 100) + 
  theme(legend.position = "none")


dev.copy(pdf, paste0(output.dir, "Frequency_of_sequenced_cells_by_Donor.pdf"))
dev.off()


# Tissue distribution
x <- table(ts.immune.seurat@meta.data$organ_tissue)
y <- x/sum(x)
y <- y*100

y <- as.data.frame(y)

ggplot(y) +
  aes(x = Var1, y = Freq, fill = Var1) +
  geom_col() +
  scale_fill_hue(direction = 1) +
  labs(y = "Freq of sequenced cells", x = "Tissue") +
  theme_minimal() +
  ylim(0, 25) + 
  theme(legend.position = "none")

dev.copy(pdf, paste0(output.dir, "Frequency_of_sequenced_cells_by_Tissue.pdf"))
dev.off()


```






